### 2023-11-17 paper-source/main.tex
Major update to code and paper. Code, reclassified stages. Paper, methods section completed and prepping to write up results.

% mnras_template.tex 
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{hyperref}
% \usepackage{enumitem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Mergers in COSMOS]{Mergers in the COSMOS Field}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. O'Ryan et al.]{
David O'Ryan$^{1}$\thanks{E-mail: d.oryan@lancaster.ac.uk}
\\
% List of institutions
$^{1}$Lancaster University \\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2023}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
   Abstract will go here.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}\label{introduction}
\noindent Galaxy interaction and merging is of fundamental importance to our current theories of galaxy formation and evolution. Our current cosmological theory, $\Lambda$CDM dictates that structure is expected to form and assemble hierarchically \citep{1978MNRAS.183..341W, 2013MNRAS.436.1765M}. Thus, understanding the underlying processes that occur in galaxy interaction and merging are paramount to understanding cosmological theories. It is well known that many underlying processes are triggered or enhanced due to galaxy interaction. The clearest result of interaction is the disruption of galactic morphology, the acceleration of disks to spheroidal systems and the formation of tidal features \citep{1972ApJ...178..623T, 1977ApJ...212..616T, 2005MNRAS.357..753G, 2009MNRAS.397..802H}, increases in star formation rates when compared to the general galaxy population \citep{1991ApJ...370L..65B, 2006ApJ...652...56B, 2014MNRAS.437.2137S, 2015ApJ...807L..16K} and, potentially, their rapid quenching \citep{2013MNRAS.430.1901H, 2023RAA....23i5026D}. There is also evidence to suggest that interaction plays a role in the activation of active galactic nuclei (AGN) \citep{2011MNRAS.418.2043E, 2015ApJ...806..219C, 2023MNRAS.523.4164H}. Therefore, with interaction having such an affect on the galaxies we see - over a long time period - understanding galaxy interaction is fundamental to understanding galaxy evolution

To better our understanding of galaxy interaction, it is often prudent to break them down into different stages of interaction. This relates to the dynamical time of the interaction, and the snapshot in the interaction we are observing. Previous works focused on simulations find that a sudden increase in star formation occurs early in the dynamical timescale of the interaction \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R}, however, confirming this definitive link observationally remains elusive \citep{2023arXiv230916531R}. The same studies could also be done for nuclear activation as well as quenching of interacting systems \citep{Papers on nuclear activation}. While we cannot yet directly measure when in the dynamical timescale star formation is enhanced from observations, we can approximate it by splitting different interactions into different stages. These stages relate directly to the dynamical timescale of the interaction. Often, they are split into four: gravitationally bounded pairs, approaching pericentre of the interaction, approaching apocentre of the interaction, and approaching coalescence.

To fully study the relationship between interaction stage and the underlying physical processes occurring we require large, statistically robust interacting galaxy samples. However, they are surprisingly difficult to create; with the first interacting galaxy catalogue (the \citet{1966ApJS...14....1A} catalogue) containing only 318 systems. More recent samples have been created either by machine learning classification \citep{2019A&A...626A..49P, 2023A&A...669A.141S}, visual classification by citizen scientists \citep{2010MNRAS.401.1043D} or by photometric parameterisation \citep{2004AJ....128..163L, 2023MNRAS.522....1N}. However, each sample has always been plagued by contamination and a loss of statistical significance when broken down into stages. It is notoriously difficult, from morphology alone, to select interacting or merging galaxies. Without redshift information it is difficult to discern if they are close together in 3D space or only by projection effects.

We elect to use the newly created interacting galaxy catalogue created by \citet{2023ApJ...948...40O} (hereafter, OR23) which found a sample of 21,926 interacting galaxy systems. This is the largest visually classified interacting galaxy catalogue to date. This sample was created using the new ESA Datalabs\footnote{ESA Datalabs: link} platform and the convolutional neural network machine 
\texttt{Zoobot} \citep{2022MNRAS.509.3966W, 2023JOSS....8.5312W}. This classified all sources in the Hubble Source Catalogue \citep{2016AJ....151..134W} from the Hubble Space Telescope (HST) archives into interacting or non-interacting. Once this was done, a rigorous process of de-duplication, contamination removal and visual inspection was conducted.

In this work, we cross match this catalogue with the Cosmic Evolutionary Survey (COSMOS) survey\footnote{DOI: \href{https://www.ipac.caltech.edu/doi/irsa/10.26131/IRSA178}{10.26131/IRSA178}}. This provides us with ancillary data for a subset of the OR23 catalogue. This ancillary data contains many galactic parameters of interest and we explore how they evolve with different stages of interaction. We define four distinct stages that occur over the course of the full dynamical timescale of interaction, from the two galaxies approaching each other to full coalescence. We primarily focus on the evolution of the stellar masses, star formation rates, surrounding environment and AGN classification from various COSMOS catalogues across the full dynamical timescale of interaction. With the photometric redshifts available from COSMOS, we also confirm a set of interacting galaxies and close pairs to investage further parameters of interest. These include the trend of projected separation of the galactic pairs on star formation rate as well as the change with mass ratio.

%, comparing to the work of \citet{2022ApJ...940....4S} a different study using COSMOS to investigate these latter relations.

This paper is laid out as follows in Section \ref{data} we explain the OR23 and COSMOS catalogues, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching selection from the different catalogues.  Section \ref{staging} describes the methods by which we split the interacting systems into their different stages, and how we identify the secondary galaxies to each identified interacting galaxy. We also discuss our processes of calculating projected separations. In Section \ref{results} we describe and discuss our results, and their implications, and compare to previous works focused on investigating different stages of interaction. Finally, Section \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

Where necessary, we use a Flat $\Lambda$CDM cosmology with $H_0$ = $70$\,km/s/Mpc and $\Omega_M = 0.3$. Hereafter in this paper, when referring to an interacting galaxy we are referring to a galaxy which has undergone one or multiple flybys by a secondary galaxy and caused tidal disturbance. A merging galaxy is the final state of these flybys, where two or more systems have or are about to coalesce to form a highly morphologically irregular system.

\section{DATA: Catalogue Matching \& Secondary Identification} \label{data}
\subsection{The O'Ryan+23 Catalogue}
\noindent We utilise the interacting galaxy catalogue created in OR23. While the full process of its creation is detailed there, we briefly describe it here. This sample of interacting galaxies was created using a convolutional neural network - \texttt{Zoobot} - to search all final product data in the \textit{Hubble} Legacy Archive. A regular convolutional neural network classifies images by extracting features from them, and using this to output a solution. \texttt{Zoobot} differs by extracting features from an image, and then reduces them to a generalised representation (essentially, a small feature vector). The power of this is that it makes \texttt{Zoobot} incredibly generalisable. This is further reinforced by training it on citizen scientist responses to the Galaxy Zoo: DECaLS \citep{2022MNRAS.509.3966W} project to every task in that project. Thus, it can answer a host of questions as well as be very easily finetuned to new datasets and classification tasks. In OR23, it was easily finetuned on \textit{Hubble} data by using classifications from all previous Galaxy Zoo projects \citep{2008MNRAS.389.1179L}.

Cutouts of each source were created as single-band \textit{F814W} images at 150 $\times$ 150 pixels across. This choice was motivated by it having the highest number of final product FITS files associated with it, while the image size was kept small due to the high number of cutouts required to be stored. Source coordinates were found in the \textit{Hubble} Source Catalogue \citep{2016AJ....151..134W}. This corresponded to just over 92 million cutouts to create and sources to classify. Due to issue of classifying interacting galaxies by morphology alone an extremely conservative prediction score from \textit{Zoobot} was required. This resulted in a very pure, but highly incomplete catalogue of 21,926 interacting galaxy systems. This is the largest catalogue of interacting galaxies directly found to date.

However, in the OR23 catalogue, only the coordinates and the \textit{Hubble} Source catalogue are provided. For ancillary data, this catalogue must be matched to different data repositories. This is not a trivial task, as a large proportion of the OR23 catalogue is known to not exist in any existing astrophysical data repository. In this work, we opt to match the OR23 catalogue to the COSMOS catalogues.

\subsection{The COSMOS2020 Catalogue}
We opt to match the OR23 catalogue with the catalogues of the COSMOS survey. Specifically, we elect to use the COSMOS2020 catalogue \citep{2022ApJS..258...11W}. This catalogue has a wealth of ancillary information for just over 1.7 million sources in a 2 square degree area of the sky. Each source found in this area has been analysed with well used astronomical software, for our purposes primarily LePhare \citep{1999MNRAS.310..540A, 2006A&A...457..841I} and EAzY \citep{2008ApJ...686.1503B}. These provide two different measurements of the parameters we wish to investigate and match with stage. The first is the stellar masses of each our cross matched galaxies and the estimated star formation rates. We opt to simply use one of these two measurements in our analysis. For the stellar mass we use the LePhare measurement and for the star formation rates we use the EAzY measurements. Using either the vice versa of this, both EAzY or both LePhare measurements led to the degeneracy's within the underlying model Spectral Energy Distribution (SED) templates being revealed. Therefore, using these two opposing parameters gave realistic approximations of these photometrically approximated values.

As stated previously, the OR23 catalogue only contained the source ID, right ascension and declination of each interacting galaxy. Therefore, we had to cross match these with the COSMOS2020 catalogue right ascension and declination measurements. To do this, we searched the COSMOS2020 catalogue for any sources within 10" of the OR23 sources. If more than one COSMOS2020 source was within this radius, we took the closest source to the OR23 coordinates as the correct source. Upon matching based on coordinates, we conducted a de-duplication of the resultant catalogue based on COSMOS2020 ID, and reassigned COSMOS2020 sources to each duplication. If no further COSMOS2020 sources were within 10" of the OR23 source, then this was considered not to be cross matched. We find that, after this cross matching and de-duplication, 3,786 of the OR23 sources exist in the COSMOS2020 catalogue.

We make two further cuts, based on the ancillary data available in the COSMOS2020 catalogue. First, there are many unphysical photometric measurements for the stellar mass or star formation rates. Therefore, we cut out these results as spurious. We only keep systems within a mass range of $6.5 \leq \log_{10} \text{M}_{*}(\text{M}_{\odot}) \leq 12.5$ and a star formation rate range of $-5 < \log_{10} \text{SFR} (M_{\odot}yr^{-1}) \leq 3.5$. Secondly, we institute a redshift cut of $z \leq 1.2$. This is to match the catalogue that will be later used to investigate the incidence of our sample of interacting galaxies with environment (detailed in Section \ref{data:environ}). Applying these cuts reduces our sample size to 3,689 interacting galaxies.

\subsection{Secondary Identification}
\noindent As stated previously, the OR23 catalogue only identified interacting galaxy systems, not individual interacting galaxies. Therefore, we must find the secondary in each interaction using the COSMOS2020 catalogue. The size of the cutouts in this catalogue were also small, and therefore, the secondary did not often appear in the cutout in that project. To find the secondaries, we apply three steps. First, a cutout surrounding the primary of each system was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles as the source of the cutouts in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels) centred and was centred on the primary galaxy. The original OR23 cutout was also displayed next to the enlarged cutout. We then searched the COSMOS2020 catalogue for all sources within the cutout. Their positions and redshifts were recorded and used to annotate the cutout.

We then visually assessed each annotated cutout, giving one of the four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift; incorrect primary assigned. Our definition of a secondary galaxy was such that the secondary had to be within the cutout we were visually assessing and within $z=\pm0.1$ of the primary redshift. We use this cutoff as this is often used when calculating environment parameters \citep[e.g][]{Baldry 2006} or defining interacting galaxies by close pairs \citep[e.g][]{Shah et al. paper from 2022 / 2023} based on the uncertainties present in using photometric redshift estimates from SED fitting. We define a null redshift as one outwith our redshift limits or NaN. A minority of the cutouts we visually assessed were found to have the incorrect primary at the centre. This was a result of our cross matching based on radius with the COSMOS catalogue.

Using these definitions, we find that of the 3,689 original systems cross-matched with COSMOS2020 2,209 could not have their secondary identified, 834 had a clear secondary, 446 could not be reliably classified as an interacting galaxy, 248 had a null redshift and 149 were the incorrect primary. Figure \ref{fig:secondary_selection} shows an example of each of our classifications. Each secondary we identify was added to our sample, increasing our sample size to 3,829.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/cutouts_ex.pdf}
    \caption{An example of each visual classification made on the cross matched sample. From left to right these are the secondary could not be identified, the primary had a clear secondary, the primary could not be reliably classified as an interacting galaxy, the redshift was null and the incorrect primary was identified in OR+23. Based on these classifications, we either add the secondary galaxies to the sample or we remove the contamination from it. These images are 30" across using the COSMOS cutout service, selecting HST/ACS tiles as the basis for the observations in the $F814W$ filter.}
    \label{fig:secondary_selection}
\end{figure*}

% Define our classifications above.
While initially surprising that the majority of our systems could not have a secondary identified, we found that it was mostly due to limitations in the COSMOS catalogue or the way in which we had conducted our secondary identification. To define a secondary galaxy, the object had to have a ID associated with it in the COSMOS2020 catalogue. However, when two systems were very close together and small enough, often they would be identified under a single COSMOS ID despite being two separate systems. The same also occurred when two systems were merging or interacting. The tidal features connecting the systems would lead to only one of them being identified in the catalogue, or when two cores appeared in one merging system. Figure \ref{fig:secondary_selection} specifically an example of two systems being close enough together that they have been identified under a single COSMOS2020 entry. Figure \ref{fig:secondaries_found} shows this disparity with the different types of interaction we observe.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/visualisation_classification.pdf}
    \caption{Where a secondary could be identified at different stages in the interaction. The reasoning for such disparity in secondaries identified is due to the relative distance each the secondary would be from the primary at each stage. For a close pair, we often found the secondary galaxy, but a minority of these were so close together that the entire system was given a single COSMOS ID. The same was true for those interacting systems approaching pericentre or merging. When the secondary was near apocentre, often it would be outwith the cutout we were using for visual classification.}
    \label{fig:secondaries_found}
\end{figure}

Those galaxies which were found to be contamination (i.e., could not be reliably classified as interacting or having a null redshift) were removed from our sample. Galaxies that could not be reliably classified as interacting were those which showed little tidal distortion and had no systems nearby with a redshift approximately that of the primary. Often, these were systems that were overlapping but at different redshifts, systems with irregular morphologies of spiral arms or clumpy galaxies. There was also many systems that were at high redshift ($z \geq 1$) where the resolution of the cutouts meant that features could not be discerned visually.

Finally, we found that there were 149 systems which had been assigned to an incorrect primary. These were systems where the interacting galaxies were clearly in the cutout but some tidal debris or some nearby system had been assigned as the primary instead. This was a result of our catalogue matching based solely on separation between the OR23 coordinates and the COSMOS2020 source coordinates. We reassign these systems to the correct COSMOS IDs and then take them through the secondary identification process.

By the end of this pipeline, we find a sample size of 3,829 interacting galaxies. We conduct a de-duplication based on the COSMOS2020 ID, which reduces the sample back down to 3,547 interacting galaxies. This shows that many of the secondaries we had identified for the different systems were already found in the OR23 catalogue. The remaining systems are each visually confirmed interacting and disturbed galaxies based on their morphology and photometric redshift as measured in the COSMOS2020 catalogue.

\subsection{Finding Additional Systems}
\noindent As a result of using visual classification to find the secondary galaxy in each interaction, we were able to also confirm other interacting systems which had not been found in the OR23 catalogue. Primarily, these extra interacting galaxies were more systems that were involved in the interactions already recorded in OR23. The pipeline described previously was only used to find a secondary for each interacting galaxy, however in many cases more galaxies were involved. Other interacting galaxies that were added here were low redshift systems which would have appeared very large in the OR23 cutout. Details and tidal features were then lost in the original classification process. By looking at the larger COSMOS2020 cutouts used here, we were able to confirm these were interacting galaxies and add them to the catalogue. The final set of extra systems we added were often those that were close pairs, but physically in 3D space. While they had little tidal disruption - what the interacting galaxies were classified on in OR23 - with the photometric redshifts to hand, we could correctly identify them here.

In total, we found an extra 841 interacting systems that we could add to our sample. Upon conducting a de-duplication of these with the sample already found, this was reduced to 634 interacting systems. This gives us a total sample size of 4,181 which we use through the rest of this work.

\section{METHOD: Environment, AGN and Interaction Stage}
In this section, we describe how we separate our sample of 4,181 interacting galaxies into separate stages. Each stage is defined to capture a different part of the interaction, and a different part of the dynamical history of the interaction. We want to compare different stages in the interaction to different physical underlying processes, such as star formation rate. Such parameters can be found in the COSMOS2020 catalogue. However, two fundamental properties we also want to compare with is that of AGN fraction evolution with time and that of environment, and the impacts this might have.

The parameters required to calculate the AGN fraction are not found in the COSMOS2020, and we therefore must cross match with other catalogues to find the required parameters. We also describe the catalogue we use here to define the density parameter around each of our sources, and define how the environment is calculated there.

\subsection{Classifying Stage of Interaction}\label{sec:staging}
The primary goal of this work is to find if a relation exists between a host of galactic parameters, underlying physical processes and the stage of the interaction. We, therefore, need to define different stages for them. We have hinted at this in the previous section with Figure \ref{fig:secondaries_found}. We break our sample down into four stages:

\begin{itemize}
    \item Stage 1: Systems which are well-separated with no morphological disturbance.
    \item Stage 2: Close pairs showing either morphological distortion or physical connection by tidal features.
    \item Stage 3: Well separated pairs with morphological disturbance or isolated galaxies with clear tidal features.
    \item Stage 4: Highly disturbed systems with two or more cores within.
\end{itemize}

\noindent Examples of each stage can be found in Figure \ref{fig:stages}. Each stage is designed to capture a different part of the interaction history of a system in the total dynamical time.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/examples-stages.pdf}
    \caption{Examples of the four stages we split our interacting galaxy sample into. Stage 1: A close pair with confirmed redshift matching. Stage 2: Two distinct systems interacting with tidal features forming. Stage 3: A tidally disturbed system with no secondary present, likely at apocentre. Stage 4: A galaxy with multiple cores while highly disturbed. At the final stage before coalescence.}
    \label{fig:stages}
\end{figure}

Stage 1 of the interaction captures the first approach of the two systems. This is the time when the two galaxies potentials are beginning to have an effect on one another. They have no morphological disturbance yet and, in our sample, are most often those galaxies with distinct disks. Hence, this stage has the highest number of if identified secondaries with every system having a secondary that can be visually classified. In this sample, we would expect no effects from the interaction to have began yet.

Stage 2 of the interaction is defined as the point where the two galaxies in the interaction are approaching or just passing the pericentre of the tidal encounter. For those galaxies before pericentre, we would expect the beginning of morphological disturbance or some irregularity in the two systems. Once the galaxies have passed pericentre, we would define it by much higher morphological disturbance, distortion and tidal debris. This requirement for morphological disturbance is how we differentiate stage 2 from stage 1 interacting galaxies. Throughout this stage of the interaction, in our sample, we see significant overlap or connection of the two systems in the form of tidal features and morphological distortion.

Stage 3 interacting galaxies are described as those interacting systems where the systems have approached apocentre from each other. In many of cases of this stage, if the galaxies have sufficient velocity, the galaxies will escape each other and not coalesce. Hence, galaxies at this stage often could not have a secondary galaxy identified in our visual assessment. We purely describe a galaxy as stage 3 based on significant tidal and morphological disturbance. This stage has the highest potential for contamination by irregular galaxies or galaxies whose morphology has been disturbed by processes other than tidal interaction. We discuss the potential impacts of this contamination, and the limitations of using visual morphology as a proxy for stage in Section \ref{sec:limitations}.

Finally, stage 4 represents the final step of a galaxy interaction. If the two galaxies do not have sufficient velocity to escape one another, then they go on to coalesce and ultimately merge. We define this stage such that the sample is of galaxies just before this coalescence and the two cores remain distinct. However, it is important to note that this stage may be contaminated by some post-merger remnants. In this context, we define a stage 4 galaxy as one with significant morphological disturbance but with multiple cores present in the disks. At this stage, the interaction is at it's most violent with complete disruption of both galactic disks. We also expect star formation rate to increase significantly in this stage. This is important to note, as high star forming at specific redshifts in the $F814W$ filter may appear like a second core within a galactic disk. At this point, we would be observing a post-merger remnant rather than the point just before coalescence. Therefore, to mitigate potential contamination in from the post-merger stage we only classify stage 4 galaxies as those with highly disrupted disks and second cores close to the primary core (and not far out in the disk).

Our four stage approach is not a new one, and many other works have utilised it to differentiate different parts of the dynamical history of a galaxy interaction \citep{}. It is, however, important to note its limitations. First, this is a simplified view of the dynamical timescale of an interaction. We have assumed that an interaction begins, the pair have enough velocity to escape or they coalesce and that is representative of the full dynamical timescale. We have not considered systems where there are multiple flybys before coalescence or systems where there are multiple members at different stages of the interaction. This is a particular concern when looking at stages 2 and 3. For instance, the effects could be enhanced or diminished over multiple flybys, and we have no way to capture this.

However, with these simplifications in our assumptions, we will investigate the parameters which are proxies for the underlying physical processes and interpret them. We will investigate the whether we see have a bias in our systems based on environment, whether we find increased AGN activity and if we observe enhancement in star formation with stage. The star formation rate is already present in the COSMOS2020 catalogue from both EAzY and LePhare. For the environment parameter and whether AGN are detected, we cross match to two further catalogues. We describe these catalogues here, our process of cross matching and the new sample sizes we have in each subsample.

\subsection{Matching to Environment Catalogue}\label{data:environ}
\noindent There is no measure of the environmental density in the COSMOS2020 catalogue. Such a measure is often calculated in numerous ways, such as the N-nearest neighbour \citep{2006MNRAS.373..469B}, different Bayesian metrics \citep{2008ApJ...674L..13C} or estimating it from Voronoi Tesselation \citep{2021inas.book...57V}. However, in this work, we utilise the existing environmental density catalogue produced by \citet{2017ApJ...837...16D}. This catalogue was created specifically for the COSMOS survey, and has a measured density for all sources with mass $\log(\frac{M}{M_\odot}) \geq 9.6$ and out to a redshift of 1.2. In this catalogue, they calculate not only the density, but also the density parameter $\delta$ and assign each source to a field, filament or cluster classification. For a full description of how they calculate the environment and density field see \citet{2015ApJ...805..121D} and \citet{2017ApJ...837...16D}, but we will briefly describe it here.

To build the density field throughout the COSMOS field, they first construct a set of overlapping redshift slices. Within each slice, a subset of the galaxies are selected such that the median of the probability distribution function (PDF) of their photometric redshift is within it. Then, from this subset, they calculate the weighted surface density within the redshift slice. The weighting is based upon the PDF of the photmetric redshift present within the redshift slice. These weights significantly reduce the effects of projection effects. They then apply a weighted adaptive kernel smoothing using a 2D Gaussian kernel whose width changes based on the found local density of galaxies. Once this density field is created, the density around the sample galaxies can simply be interpolated across the density field based on the angular position and the redshift slice the sample galaxy is in (based on its photometric redshift PDF).

The result of this process, and the cuts defined previously, is a catalogue of $\approx$45,000 galaxies with their densities accurately measured. As advised with the catalogue, we remove any sources which are flagged as uncertain. This is recorded in the catalogue. This leaves us with $\approx$39,000 sources with which to cross match the OR23 sample. We apply the same cuts to our sample as applied in \citet{2017ApJ...837...16D}, and only consider those systems with a mass $\log(\frac{M}{M_\odot}) \geq 9.6$. To cross match with our sample, we simply utilise the COSMOS2015\_ID which is recorded in the COSMOS2020 catalogue which corresponds to the ID in the \citet{2017ApJ...837...16D} catalogue. We find that every source in our COSMOS2020 sample has a COSMOS2015 ID associated with it, and therefore we fully match the reduced catalogue.

Upon applying the mass cut to our sample, we find that we are matching 2,800 sources to the \citet{2017ApJ...837...16D} catalogue. Upon matching based upon the COSMOS2015 ID, we find that 628 sources in our sample do not exist in the environment catalogue. This reduces our sample to 2,172 galaxies with confirmed and reliable environment density measurements.

\subsection{Classifying AGN}
\noindent We will also investigate the affect of interaction stage upon the AGN activity within the galaxies involved. Unfortunately, the outputs of EAzY and LePhare do not provide us with photometric estimates of the line fluxes needed to define an AGN on a Baldwin, Philips and Terlevich (BPT) diagram \citep{}. Therefore, we cannot directly identify if a galaxy is only star forming, has nuclear activity or is a composite. So, we turn to cross-matching our sample to three different catalogues which contain the required information. We use the Chandra COSMOS Legacy Survey Multiwavelength Catalogue \citep{2016ApJ...817...34M} and the COSMOS VLA 3GHz survey \citep{2017A&A...602A...6S, 2017A&A...602A...3D}. In these catalogues an AGN, star forming or composite classification is given within the catalogue.

Again, our cross match process is very much the same as previously described. We use the previously identified COSMOS2020 source coordinates and conduct a radial search of 10". If more than one source is within this radius, we match on the closest source. The two catalogues derived from the COSMOS survey do not contain IDs which overlap with the COSMOS2020 catalogue, therefore we must do it based on coordinates. We conduct our matching in serial, starting from the \citet{2017A&A...602A...6S} catalogue and then to the Chandra survey. At every step, if we find a match in the relevant catalogue, we remove it from subsequent searches in other catalogues and take the first classification as the correct one.

Applying our matching criteria, we find 1,059 classifications in the VLA 3GHz survey and 155 in the Chandra survey. These were split into 812 star forming galaxies and 402 AGN. We also investigate cross matching with the MPA-JHU catalogue \citep{2003MNRAS.341...33K, 2004MNRAS.351.1151B, 2007ApJS..173..267S}, however, found that all matches were already represented by the VLA and Chandra surveys. Therefore, we only utilise these two catalogues in our sample.

While the ratio of AGN to star forming galaxies in our sample seems large compared to other works, it is important to note that this is a result of limited matching between the catalogues. Of the 4,182 galaxies in our sample, only 1,214 appeared at all in either the VLA or Chandra catalogues. Any galaxy which did not appear in these catalogues, we discard as unclassified. Therefore, we require these catalogues to specifically show us if a galaxy is star forming or not before we classify it as such. This is why our number of star forming galaxies is lower than expected. We discuss this limitation in section \ref{sec:limitations} and the effect it may have on our results of AGN fraction evolution with stage.

\section{Results \& Discussion}\label{results}
\noindent With our samples selected, and numbers throughout known, we can begin to investigate teh change in multiple parameters with stage of the interaction. As stated previously, the four stages of interaction are designed to capture the main different parts of galaxy interaction. First, we show the results of breaking down our sample into stages with relation to the star formation and mass of our sample. We investigate the differences between these distributions and interpret what underlying physical processes could be occurring to show these differences. We show that the results we have found could not be due to other effects, like different environments affecting our selection etc. We then this up by discussing the changing AGN sample that we see, discuss selection effects and discuss the potential underlying processes responsible for our results.

Finally, to put this in more context with the literature, we break our sample down further into correctly identified pairs and mergers. We use this subsample to explore the relationship between the galactic parameters previously discussed, and the mass ratio and projected separations of the systems. We show that this agrees with previous works, and how this lends legitimacy to our results with different stages of interaction.

\subsection{Star Formation Rate with Stage}

\subsection{Controlling for Galactic Environment}

\subsection{Nuclear Activity with Stage}

\subsection{Star Formation and Mass Ratio}

\subsection{Star Formation and Projected Separation}

\subsection{Comparison to Other Works}

\section{Limitations of Approach}\label{sec:limitations}
\subsection{Redshift Effects on Visual Classification}
\noindent An important limitation of our visual approach is related to the effect of tidal features out to higher redshifts. As the redshift increases, surface brightness dimming of the systems may lead to a lack of visual confirmation in the tidal features of different objects. Thus, if we were heavily affected by this in our redshift range ($0.0 \leq z \leq 1.2$), we would see distinct signs of it in our selection function of our mergers. For instance, if all of our stage 4 classifications were actually high redshift stage 3 galaxies, we would observe this in our selection function. Therefore, Figure \ref{fig:redshift_selection} shows our selection functions across our redshift range for each defined stage in our interaction.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/stage-selection.pdf}
    \caption{Redshift vs Mass distribution for each stage of interaction we have defined. This is important as we use tidal distortion and the existence as tidal features as a fundamental for our classification methodology. As shown here, the distribution of systems across redshift and mass is consistent for all stages in our sample. Therefore, we are likely not affected by this in our analysis.}
    \label{fig:redshift_selection}
\end{figure}

As shown, the distribution of masses across redshift that we find in our sample is reasonably consistent with each other. This means we are likely not affected by surface brightness dimming across our sample and our redshift range.

\subsection{Selection Effects in AGN Sample}

\subsection{Limitations of Visual Classification}

\section{Conclusions}\label{conclusion}
 - Restate the Results \\
 - What's the next steps of this project?
    - Hint at what I would do in a Fellowship.

\section*{Acknowledgements}
The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

This work utilised data from the Cosmic Evolution Survey (COSMOS) which has a digital object identifier of 10.26131/IRSA178.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:
\bibliographystyle{mnras}
\bibliography{references} % if your bibtex file is called example.bib

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Needed?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex