### 2023-11-29 paper-source/main.tex
Major update to finding weighted AGN fraction. Also added to paper and lots of projected separation stuff. Paper nearly there, just gotta write a discussion and conclusion.

% mnras_template.tex 
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{hyperref}
\usepackage{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Mergers in COSMOS]{Mergers in the COSMOS Field}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. O'Ryan et al.]{
David O'Ryan$^{1}$\thanks{E-mail: d.oryan@lancaster.ac.uk}
\\
% List of institutions
$^{1}$Lancaster University \\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2023}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
   Abstract will go here.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}\label{introduction}
\noindent Galaxy interaction and merging is of fundamental importance to our current theories of galaxy formation and evolution. Our current cosmological theory, $\Lambda$CDM dictates that structure is expected to form and assemble hierarchically \citep{1978MNRAS.183..341W, 2013MNRAS.436.1765M}. Thus, understanding the underlying processes that occur in galaxy interaction and merging are paramount to understanding cosmological theories. It is well known that many underlying processes are triggered or enhanced due to galaxy interaction. The clearest result of interaction is the disruption of galactic morphology, the acceleration of disks to spheroidal systems and the formation of tidal features \citep{1972ApJ...178..623T, 1977ApJ...212..616T, 2005MNRAS.357..753G, 2009MNRAS.397..802H}, increases in star formation rates when compared to the general galaxy population \citep{1991ApJ...370L..65B, 2006ApJ...652...56B, 2014MNRAS.437.2137S, 2015ApJ...807L..16K} and, potentially, their rapid quenching \citep{2013MNRAS.430.1901H, 2023RAA....23i5026D}. There is also evidence to suggest that interaction plays a role in the activation of active galactic nuclei (AGN) \citep{2011MNRAS.418.2043E, 2015ApJ...806..219C, 2023MNRAS.523.4164H}. Therefore, with interaction having such an affect on the galaxies we see - over a long time period - understanding galaxy interaction is fundamental to understanding galaxy evolution

To better our understanding of galaxy interaction, it is often prudent to break them down into different stages of interaction. This relates to the dynamical time of the interaction, and the snapshot in the interaction we are observing. Previous works focused on simulations find that a sudden increase in star formation occurs early in the dynamical timescale of the interaction \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R}, however, confirming this definitive link observationally remains elusive \citep{2023arXiv230916531R}. The same studies could also be done for nuclear activation as well as quenching of interacting systems \citep{2011MNRAS.418.2043E, 2018PASJ...70S..37G, 2023ApJ...942..107S}. While we cannot yet directly measure when in the dynamical timescale star formation is enhanced from observations, we can approximate it by splitting different interactions into different stages. These stages relate directly to the dynamical timescale of the interaction. Often, they are split into four: gravitationally bounded pairs, approaching pericentre of the interaction, approaching apocentre of the interaction, and approaching coalescence.

To fully study the relationship between interaction stage and the underlying physical processes occurring we require large, statistically robust interacting galaxy samples. However, they are surprisingly difficult to create; with the first interacting galaxy catalogue (the \citet{1966ApJS...14....1A} catalogue) containing only 318 systems. More recent samples have been created either by machine learning classification \citep{2019A&A...626A..49P, 2023A&A...669A.141S}, visual classification by citizen scientists \citep{2010MNRAS.401.1043D} or by photometric parameterisation \citep{2004AJ....128..163L, 2023MNRAS.522....1N}. However, each sample has always been plagued by contamination and a loss of statistical significance when broken down into stages. It is notoriously difficult, from morphology alone, to select interacting or merging galaxies. Without redshift information it is difficult to discern if they are close together in 3D space or only by projection effects.

We elect to use the newly created interacting galaxy catalogue created by \citet{2023ApJ...948...40O} (hereafter, OR23) which found a sample of 21,926 interacting galaxy systems. This is the largest visually classified interacting galaxy catalogue to date. This sample was created using the new ESA Datalabs\footnote{ESA Datalabs: link} platform and the convolutional neural network machine 
\texttt{Zoobot} \citep{2022MNRAS.509.3966W, 2023JOSS....8.5312W}. This classified all sources in the Hubble Source Catalogue \citep{2016AJ....151..134W} from the Hubble Space Telescope (HST) archives into interacting or non-interacting. Once this was done, a rigorous process of de-duplication, contamination removal and visual inspection was conducted.

In this work, we cross match this catalogue with the Cosmic Evolutionary Survey (COSMOS) survey\footnote{DOI: \href{https://www.ipac.caltech.edu/doi/irsa/10.26131/IRSA178}{10.26131/IRSA178}}. This provides us with ancillary data for a subset of the OR23 catalogue. This ancillary data contains many galactic parameters of interest and we explore how they evolve with different stages of interaction. We define four distinct stages that occur over the course of the full dynamical timescale of interaction, from the two galaxies approaching each other to full coalescence. We primarily focus on the evolution of the stellar masses, star formation rates, surrounding environment and AGN classification from various COSMOS catalogues across the full dynamical timescale of interaction. With the photometric redshifts available from COSMOS, we also confirm a set of interacting galaxies and close pairs to investage further parameters of interest. These include the trend of projected separation of the galactic pairs on star formation rate as well as the change with mass ratio.

%, comparing to the work of \citet{2022ApJ...940....4S} a different study using COSMOS to investigate these latter relations.

This paper is laid out as follows in Section \ref{data} we explain the OR23 and COSMOS catalogues, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching selection from the different catalogues.  Section \ref{method} describes the methods by which we split the interacting systems into their different stages, and how we identify the secondary galaxies to each identified interacting galaxy. We also discuss our processes of calculating projected separations. In Section \ref{results} we describe and discuss our results, and their implications, and compare to previous works focused on investigating different stages of interaction. Finally, Section \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

Where necessary, we use a Flat $\Lambda$CDM cosmology with $H_0$ = $70$\,km/s/Mpc and $\Omega_M = 0.3$. Hereafter in this paper, when referring to an interacting galaxy we are referring to a galaxy which has undergone one or multiple flybys by a secondary galaxy and caused tidal disturbance. A merging galaxy is the final state of these flybys, where two or more systems have or are about to coalesce to form a highly morphologically irregular system.

\section{DATA: Catalogue Matching \& Secondary Identification} \label{data}
\subsection{The O'Ryan+23 Catalogue}
\noindent We utilise the interacting galaxy catalogue created in OR23. While the full process of its creation is detailed there, we briefly describe it here. This sample of interacting galaxies was created using a convolutional neural network - \texttt{Zoobot} - to search all final product data in the \textit{Hubble} Legacy Archive. A regular convolutional neural network classifies images by extracting features from them, and using this to output a solution. \texttt{Zoobot} differs by extracting features from an image, and then reduces them to a generalised representation (essentially, a small feature vector). The power of this is that it makes \texttt{Zoobot} incredibly generalisable. This is further reinforced by training it on citizen scientist responses to the Galaxy Zoo: DECaLS \citep{2022MNRAS.509.3966W} project to every task in that project. Thus, it can answer a host of questions as well as be very easily finetuned to new datasets and classification tasks. In OR23, it was easily finetuned on \textit{Hubble} data by using classifications from all previous Galaxy Zoo projects \citep{2008MNRAS.389.1179L}.

Cutouts of each source were created as single-band \textit{F814W} images at 150 $\times$ 150 pixels across. This choice was motivated by it having the highest number of final product FITS files associated with it, while the image size was kept small due to the high number of cutouts required to be stored. Source coordinates were found in the \textit{Hubble} Source Catalogue \citep{2016AJ....151..134W}. This corresponded to just over 92 million cutouts to create and sources to classify. Due to issue of classifying interacting galaxies by morphology alone an extremely conservative prediction score from \textit{Zoobot} was required. This resulted in a very pure, but highly incomplete catalogue of 21,926 interacting galaxy systems. This is the largest catalogue of interacting galaxies directly found to date.

However, in the OR23 catalogue, only the coordinates and the \textit{Hubble} Source catalogue are provided. For ancillary data, this catalogue must be matched to different data repositories. This is not a trivial task, as a large proportion of the OR23 catalogue is known to not exist in any existing astrophysical data repository. In this work, we opt to match the OR23 catalogue to the COSMOS catalogues.

\subsection{The COSMOS2020 Catalogue}
We opt to match the OR23 catalogue with the catalogues of the COSMOS survey. Specifically, we elect to use the COSMOS2020 catalogue \citep{2022ApJS..258...11W}. This catalogue has a wealth of ancillary information for just over 1.7 million sources in a 2 square degree area of the sky. Each source found in this area has been analysed with well used astronomical software, for our purposes primarily LePhare \citep{1999MNRAS.310..540A, 2006A&A...457..841I} and EAzY \citep{2008ApJ...686.1503B}. These provide two different measurements of the parameters we wish to investigate and match with stage. The first is the stellar masses of each our cross matched galaxies and the estimated star formation rates. We opt to simply use one of these two measurements in our analysis. For the stellar mass we use the LePhare measurement and for the star formation rates we use the EAzY measurements. Using either the vice versa of this, both EAzY or both LePhare measurements led to the degeneracy's within the underlying model Spectral Energy Distribution (SED) templates being revealed. Therefore, using these two opposing parameters gave realistic approximations of these photometrically approximated values.

As stated previously, the OR23 catalogue only contained the source ID, right ascension and declination of each interacting galaxy. Therefore, we had to cross match these with the COSMOS2020 catalogue right ascension and declination measurements. To do this, we searched the COSMOS2020 catalogue for any sources within 10" of the OR23 sources. If more than one COSMOS2020 source was within this radius, we took the closest source to the OR23 coordinates as the correct source. Upon matching based on coordinates, we conducted a de-duplication of the resultant catalogue based on COSMOS2020 ID, and reassigned COSMOS2020 sources to each duplication. If no further COSMOS2020 sources were within 10" of the OR23 source, then this was considered not to be cross matched. We find that, after this cross matching and de-duplication, 3,786 of the OR23 sources exist in the COSMOS2020 catalogue.

We make two further cuts, based on the ancillary data available in the COSMOS2020 catalogue. First, there are many unphysical photometric measurements for the stellar mass or star formation rates. Therefore, we cut out these results as spurious. We only keep systems within a mass range of $6.5 \leq \log_{10} \text{M}_{*}(\text{M}_{\odot}) \leq 12.5$ and a star formation rate range of $-5 < \log_{10} \text{SFR} (M_{\odot}yr^{-1}) \leq 3.5$. Secondly, we institute a redshift cut of $z \leq 1.2$. This is to match the catalogue that will be later used to investigate the incidence of our sample of interacting galaxies with environment (detailed in Section \ref{data:environ}). Applying these cuts reduces our sample size to 3,689 interacting galaxies.

\subsection{Secondary Identification}\label{sec:sec-ident}
\noindent As stated previously, the OR23 catalogue only identified interacting galaxy systems, not individual interacting galaxies. Therefore, we must find the secondary in each interaction using the COSMOS2020 catalogue. The size of the cutouts in this catalogue were also small, and therefore, the secondary did not often appear in the cutout in that project. To find the secondaries, we apply three steps. First, a cutout surrounding the primary of each system was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles as the source of the cutouts in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels) centred and was centred on the primary galaxy. The original OR23 cutout was also displayed next to the enlarged cutout. We then searched the COSMOS2020 catalogue for all sources within the cutout. Their positions and redshifts were recorded and used to annotate the cutout.

We then visually assessed each annotated cutout, giving one of the four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift; incorrect primary assigned. Our definition of a secondary galaxy was such that the secondary had to be within the cutout we were visually assessing and within $z=\pm0.1$ of the primary redshift. We use this cutoff as this is often used when calculating environment parameters \citep[e.g][]{2006MNRAS.373..469B} or defining interacting galaxies by close pairs \citep[e.g][]{2022ApJ...940....4S} based on the uncertainties present in using photometric redshift estimates from SED fitting. We define a null redshift as one outwith our redshift limits or NaN. A minority of the cutouts we visually assessed were found to have the incorrect primary at the centre. This was a result of our cross matching based on radius with the COSMOS catalogue.

Using these definitions, we find that of the 3,689 original systems cross-matched with COSMOS2020 2,209 could not have their secondary identified, 834 had a clear secondary, 446 could not be reliably classified as an interacting galaxy, 248 had a null redshift and 149 were the incorrect primary. Figure \ref{fig:secondary_selection} shows an example of each of our classifications. Each secondary we identify was added to our sample, increasing our sample size to 3,829.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/cutouts_ex.pdf}
    \caption{An example of each visual classification made on the cross matched sample. From left to right these are the secondary could not be identified, the primary had a clear secondary, the primary could not be reliably classified as an interacting galaxy, the redshift was null and the incorrect primary was identified in OR+23. Based on these classifications, we either add the secondary galaxies to the sample or we remove the contamination from it. These images are 30" across using the COSMOS cutout service, selecting HST/ACS tiles as the basis for the observations in the $F814W$ filter.}
    \label{fig:secondary_selection}
\end{figure*}

% Define our classifications above.
While initially surprising that the majority of our systems could not have a secondary identified, we found that it was mostly due to limitations in the COSMOS catalogue or the way in which we had conducted our secondary identification. To define a secondary galaxy, the object had to have a ID associated with it in the COSMOS2020 catalogue. However, when two systems were very close together and small enough, often they would be identified under a single COSMOS ID despite being two separate systems. The same also occurred when two systems were merging or interacting. The tidal features connecting the systems would lead to only one of them being identified in the catalogue, or when two cores appeared in one merging system. Figure \ref{fig:secondary_selection} specifically an example of two systems being close enough together that they have been identified under a single COSMOS2020 entry. Figure \ref{fig:secondaries_found} shows this disparity with the different types of interaction we observe.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/visualisation_classification.pdf}
    \caption{Where a secondary could be identified at different stages in the interaction. The reasoning for such disparity in secondaries identified is due to the relative distance each the secondary would be from the primary at each stage. For a close pair, we often found the secondary galaxy, but a minority of these were so close together that the entire system was given a single COSMOS ID. The same was true for those interacting systems approaching pericentre or merging. When the secondary was near apocentre, often it would be outwith the cutout we were using for visual classification.}
    \label{fig:secondaries_found}
\end{figure}

Those galaxies which were found to be contamination (i.e., could not be reliably classified as interacting or having a null redshift) were removed from our sample. Galaxies that could not be reliably classified as interacting were those which showed little tidal distortion and had no systems nearby with a redshift approximately that of the primary. Often, these were systems that were overlapping but at different redshifts, systems with irregular morphologies of spiral arms or clumpy galaxies. There was also many systems that were at high redshift ($z \geq 1$) where the resolution of the cutouts meant that features could not be discerned visually.

Finally, we found that there were 149 systems which had been assigned to an incorrect primary. These were systems where the interacting galaxies were clearly in the cutout but some tidal debris or some nearby system had been assigned as the primary instead. This was a result of our catalogue matching based solely on separation between the OR23 coordinates and the COSMOS2020 source coordinates. We reassign these systems to the correct COSMOS IDs and then take them through the secondary identification process.

By the end of this pipeline, we find a sample size of 3,829 interacting galaxies. We conduct a de-duplication based on the COSMOS2020 ID, which reduces the sample back down to 3,547 interacting galaxies. This shows that many of the secondaries we had identified for the different systems were already found in the OR23 catalogue. The remaining systems are each visually confirmed interacting and disturbed galaxies based on their morphology and photometric redshift as measured in the COSMOS2020 catalogue.

Each of the galaxies that have a confirmed secondary will be used in measuring effects of projected separation with star formation and AGN fraction. Therefore, we must define a control sample by which to compare to. We find these control galaxies from the COSMOS2020 catalogue. All galaxies within 0.01 dex of our sub-samples' stellar mass are selected from the catalogue, and within a $\pm$0.01 redshift slice. We then order the control galaxies to be whichever system is furthest from our target. Each control galaxy is then visually confirmed to be non-interacting and in isolation. We also confirm that it does not already exist in the catalogue, and ensure there is no duplication in the control sample. Figure \ref{fig:matched-distributions} shows the mass distribution of our paired sub-sample of galaxies and their control, showing a closely similar mass distribution. With the previously defined cuts, we find a control galaxy for all but one of our galaxies with a confirmed secondary.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/mass-matching-pairs.pdf}
    \caption{The mass distribution of the paired interacting galaxy sample and the control sample. Both primary and secondary galaxies are within this distribution. From our sample of 4,135 interacting galaxies morphologically identified, 960 are confirmed galaxy pairs. This means that the secondary in the interacting galaxy system has been identified from morphology and photometric redshifts.}
    \label{fig:matched-distributions}
\end{figure}

\subsection{Finding Additional Systems}
\noindent As a result of using visual classification to find the secondary galaxy in each interaction, we were able to also confirm other interacting systems which had not been found in the OR23 catalogue. Primarily, these extra interacting galaxies were more systems that were involved in the interactions already recorded in OR23. The pipeline described previously was only used to find a secondary for each interacting galaxy, however in many cases more galaxies were involved. Other interacting galaxies that were added here were low redshift systems which would have appeared very large in the OR23 cutout. Details and tidal features were then lost in the original classification process. By looking at the larger COSMOS2020 cutouts used here, we were able to confirm these were interacting galaxies and add them to the catalogue. The final set of extra systems we added were often those that were close pairs, but physically in 3D space. While they had little tidal disruption - what the interacting galaxies were classified on in OR23 - with the photometric redshifts to hand, we could correctly identify them here.

In total, we found an extra 841 interacting systems that we could add to our sample. Upon conducting a de-duplication of these with the sample already found, this was reduced to 634 interacting systems. This gives us a total sample size of 4,181 which we use through the rest of this work.

\subsection{Redshift Effects on Visual Classification}
\noindent There are two limits when using visual morphology to find interacting galaxies. The first is the limitation of being able to characterise tidal features with increasing redshift. As the redshift increases, surface brightness dimming of the systems may lead to a lack of visual confirmation in the tidal features of different objects. Thus, if we are heavily affected by this in our redshift range ($0.0 \leq z \leq 1.2$), we would see distinct signs of it in our selection function of our mergers. For instance, if all of our stage 4 classifications were actually high redshift stage 3 galaxies, we would observe this in our selection function. Therefore, Figure \ref{fig:redshift_selection} shows our selection functions across our redshift range for each defined stage in our interaction. As shown, the distribution of masses across redshift that we find in our sample is reasonably consistent with each other. This means we are likely not affected by surface brightness dimming across our sample and our redshift range.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/stage-selection.pdf}
    \caption{Redshift vs Mass distribution for each stage of interaction we have defined. This is important as we use tidal distortion and the existence as tidal features as a fundamental for our classification methodology. As shown here, the distribution of systems across redshift and mass is consistent for all stages in our sample. Therefore, we are likely not affected by this in our analysis.}
    \label{fig:redshift_selection}
\end{figure}

The second issue that could affect our sample is that of contamination by close pairs by projection effects rather than by galaxies which are physically close together. For the majority of these systems, we are limited to only having their calculated photometric redshift from the COSMSO2020 catalogue and not their well-constrained spectroscopic redshift. As we are conducting visual classification, this issue is rather limited. The criteria on all of our sources besides stage 1 is that they must be morphologically disturbed. Each source is confirmed to have some kind of tidal feature f=present in the image, even if a secondary is not present in the image. There are, always, of course other galaxies in each image and there is no guarantee that the calcualted photometric redshift is correct. WE often find galaxies which are clearly interacting or paired, but with the photometric wildly different. In these cases, we defer to the redshift and only record one of the galaxies in our sample. This approach, however, means we must be careful to investigate the effects of environment on our galaxies as disturbance by, say, ram pressure stripping can appear like tidal features.

\section{METHOD: Environment, AGN and Interaction Stage} \label{method}
In this section, we describe how we separate our sample of 4,181 interacting galaxies into separate stages. Each stage is defined to capture a different part of the interaction, and a different part of the dynamical history of the interaction. We want to compare different stages in the interaction to different physical underlying processes, such as star formation rate. Such parameters can be found in the COSMOS2020 catalogue. However, two fundamental properties we also want to compare with is that of AGN fraction evolution with time and that of environment, and the impacts this might have.

The parameters required to calculate the AGN fraction are not found in the COSMOS2020, and we therefore must cross match with other catalogues to find the required parameters. We also describe the catalogue we use here to define the density parameter around each of our sources, and define how the environment is calculated there.

\subsection{Classifying Stage of Interaction}\label{sec:staging}
The primary goal of this work is to find if a relation exists between a host of galactic parameters, underlying physical processes and the stage of the interaction. We, therefore, need to define different stages for them. We have hinted at this in the previous section with Figure \ref{fig:secondaries_found}. We break our sample down into four stages:

\begin{itemize}
    \item Stage 1: Systems which are well-separated with little to no morphological disturbance.
    \item Stage 2: Close pairs showing either morphological distortion or physical connection by tidal features.
    \item Stage 3: Well separated pairs with morphological disturbance or isolated galaxies with clear tidal features.
    \item Stage 4: Highly disturbed systems with two or more cores within.
\end{itemize}

\noindent Examples of each stage can be found in Figure \ref{fig:stages}. Each stage is designed to capture a different part of the interaction history of a system in the total dynamical time.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/examples-stages.pdf}
    \caption{Examples of the four stages we split our interacting galaxy sample into. Stage 1: A close pair with confirmed redshift matching. Stage 2: Two distinct systems interacting with tidal features forming. Stage 3: A tidally disturbed system with no secondary present, likely at apocentre. Stage 4: A galaxy with multiple cores while highly disturbed. At the final stage before coalescence.}
    \label{fig:stages}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/projected-seps.pdf}
    \caption{The projected separations of the confirmed galaxy pairs in our sample. This confers with other works the definition of our different stages. Stage 1 can be at any projected separation, however, we have visually confirmed that these sources are not morphologically disturbed. Stage 2 is dominated by systems with small projected separation as, by definition, they must be morphologically linked or overlapping. Those systems at larger separation are very large systems whose morphology categorise them as stage 2. Finally, stage 3 galaxies are those which are visually confirmed to be fully morphologically separated and tidally disturbed. The bulk of these lie in a range of 25 - 100kpc in projected separation from each other. There is some overlap between stage 2 and stage 3 in projected separation, as their visual classification is also dependent on the system size.}
    \label{fig:proj-seps}
\end{figure*}

Stage 1 of the interaction captures the first approach of the two systems. This is the time when the two galaxies potentials are beginning to have an effect on one another. They have no morphological disturbance yet and, in our sample, are most often those galaxies with distinct disks. Hence, this stage has the highest number of if identified secondaries with every system having a secondary that can be visually classified. In this sample, we would expect no effects from the interaction to have began yet.

Stage 2 of the interaction is defined as the point where the two galaxies in the interaction are approaching or just passing the pericentre of the tidal encounter. For those galaxies before pericentre, we would expect the beginning of morphological disturbance or some irregularity in the two systems. Once the galaxies have passed pericentre, we would define it by much higher morphological disturbance, distortion and tidal debris. This requirement for morphological disturbance is how we differentiate stage 2 from stage 1 interacting galaxies. Throughout this stage of the interaction, in our sample, we see significant overlap or connection of the two systems in the form of tidal features and morphological distortion.

Stage 3 interacting galaxies are described as those interacting systems where the systems have approached apocentre from each other. In many of cases of this stage, if the galaxies have sufficient velocity, the galaxies will escape each other and not coalesce. Hence, galaxies at this stage often could not have a secondary galaxy identified in our visual assessment. We purely describe a galaxy as stage 3 based on significant tidal and morphological disturbance.

Finally, stage 4 represents the final step of a galaxy interaction. If the two galaxies do not have sufficient velocity to escape one another, then they go on to coalesce and ultimately merge. We define this stage such that the sample is of galaxies just before this coalescence and the two cores remain distinct. However, it is important to note that this stage may be contaminated by some post-merger remnants. In this context, we define a stage 4 galaxy as one with significant morphological disturbance but with multiple cores present in the disks. At this stage, the interaction is at it's most violent with complete disruption of both galactic disks. We also expect star formation rate to increase significantly in this stage. This is important to note, as high star forming at specific redshifts in the $F814W$ filter may appear like a second core within a galactic disk. At this point, we would be observing a post-merger remnant rather than the point just before coalescence. Therefore, to mitigate potential contamination in from the post-merger stage we only classify stage 4 galaxies as those with highly disrupted disks and second cores close to the primary core (and not far out in the disk).

Our four stage approach is not a new one, and many other works have utilised it to differentiate different parts of the dynamical history of a galaxy interaction \citep{2022ApJ...937...97C, 2023ApJ...952..122G}. It is, however, important to note its limitations. First, this is a simplified view of the dynamical timescale of an interaction. We have assumed that an interaction begins, the pair have enough velocity to escape or they coalesce and that is representative of the full dynamical timescale. We have not considered systems where there are multiple flybys before coalescence or systems where there are multiple members at different stages of the interaction. This is a particular concern when looking at stages 2 and 3. For instance, the effects could be enhanced or diminished over multiple flybys, and we have no way to capture this.

We relate our classification of stage to projected separations between systems, more often used in the literature, by using our confirmed galaxy pair subsample. Figure \ref{fig:proj-seps} shows the stage classification with projected separation between the two pairs. The projected separation is measured by taking the average of the best fit photometric redshifts between the two galaxies, and converting their angular separation to a physical one. The most dinstinct separation in stages in between stage 2 and stage 3. Here, we see that stage 2 is dominated by systems with a projected separation $<$35kpc while stage 3 is dominated $25 \leq s_{proj} \leq 100$ kpc. This is to be expected from our definitions of stage. Stage 2 is visually classified as systems which are highly morphologically disturbed, while either being morphologically linked to each other or overlapping. Those galaxy pairs with large projected separations are pairs which are very large in angular size, while still overlapping or morphologically linked. If this criteria is not met, then the system becomes a stage 3 interaction where the two galaxies in the pair are completely dinstinct. There is some overlap between the projected separations of stage 2 and stage 3 as this is somewhat dependent on the angular size of the two systems involved in the interaction.

We will investigate the parameters which are proxies for the underlying physical processes of our sample and interpret them. We will investigate whether we see have a bias in our systems based on environment, whether we find increased AGN activity and if we observe enhancement in star formation with stage. The star formation rate is already present in the COSMOS2020 catalogue from both EAzY and LePhare. For the environment parameter and whether AGN are detected, we cross match to two further catalogues. We describe these catalogues here, our process of cross matching and the new sample sizes we have in each subsample.

\subsection{Matching to Environment Catalogue}\label{data:environ}
\noindent There is no measure of the environmental density in the COSMOS2020 catalogue. Such a measure is often calculated in numerous ways, such as the N-nearest neighbour \citep{2006MNRAS.373..469B}, different Bayesian metrics \citep{2008ApJ...674L..13C} or estimating it from Voronoi Tesselation \citep{2021inas.book...57V}. However, in this work, we utilise the existing environmental density catalogue produced by \citet{2017ApJ...837...16D}. This catalogue was created specifically for the COSMOS survey, and has a measured density for all sources with mass $\log(\frac{M}{M_\odot}) \geq 9.6$ and out to a redshift of 1.2. In this catalogue, they calculate not only the density, but also the density parameter $\delta$ and assign each source to a field, filament or cluster classification. For a full description of how they calculate the environment and density field see \citet{2015ApJ...805..121D} and \citet{2017ApJ...837...16D}, but we will briefly describe it here.

To build the density field throughout the COSMOS field, they first construct a set of overlapping redshift slices. Within each slice, a subset of the galaxies are selected such that the median of the probability distribution function (PDF) of their photometric redshift is within it. Then, from this subset, they calculate the weighted surface density within the redshift slice. The weighting is based upon the PDF of the photmetric redshift present within the redshift slice. These weights significantly reduce the effects of projection effects. They then apply a weighted adaptive kernel smoothing using a 2D Gaussian kernel whose width changes based on the found local density of galaxies. Once this density field is created, the density around the sample galaxies can simply be interpolated across the density field based on the angular position and the redshift slice the sample galaxy is in (based on its photometric redshift PDF).

The result of this process, and the cuts defined previously, is a catalogue of $\approx$45,000 galaxies with their densities accurately measured. As advised with the catalogue, we remove any sources which are flagged as uncertain. This is recorded in the catalogue. This leaves us with $\approx$39,000 sources with which to cross match the OR23 sample. We apply the same cuts to our sample as applied in \citet{2017ApJ...837...16D}, and only consider those systems with a mass $\log(\frac{M}{M_\odot}) \geq 9.6$. To cross match with our sample, we simply utilise the COSMOS2015\_ID which is recorded in the COSMOS2020 catalogue which corresponds to the ID in the \citet{2017ApJ...837...16D} catalogue. We find that every source in our COSMOS2020 sample has a COSMOS2015 ID associated with it, and therefore we fully match the reduced catalogue.

Upon applying the mass cut to our sample, we find that we are matching 2,800 sources to the \citet{2017ApJ...837...16D} catalogue. Upon matching based upon the COSMOS2015 ID, we find that 628 sources in our sample do not exist in the environment catalogue. This reduces our sample to 2,172 galaxies with confirmed and reliable environment density measurements.

\subsection{Classifying AGN}\label{sec:agn-clsf}
\noindent We will also investigate the affect of interaction stage upon the AGN activity within the galaxies involved. Unfortunately, the outputs of EAzY and LePhare do not provide us with photometric estimates of the line fluxes needed to define an AGN on a Baldwin, Philips and Terlevich (BPT) diagram. Therefore, we cannot directly identify if a galaxy is only star forming, has nuclear activity or is a composite. So, we turn to cross-matching our sample to three different catalogues which contain the required information. We use the Chandra COSMOS Legacy Survey Multiwavelength Catalogue \citep{2016ApJ...817...34M} and the COSMOS VLA 3GHz survey \citep{2017A&A...602A...6S, 2017A&A...602A...3D}. Each source has an AGN, star forming or composite classification is within the catalogue.

% Need a small paragraph here about these two catalogues and their differences.

Again, our cross match process is very much the same as previously described. We use the previously identified COSMOS2020 source coordinates and conduct a radial search of 10". If more than one source is within this radius, we match on the closest source. The two catalogues derived from the COSMOS survey do not contain IDs which overlap with the COSMOS2020 catalogue, therefore we must do it based on coordinates. We conduct our matching in serial, starting from the \citet{2017A&A...602A...6S} catalogue and then to the Chandra survey. At every step, if we find a match in the relevant catalogue, we remove it from subsequent searches in other catalogues and take the first classification as the correct one.

Applying our matching criteria, we find 1,059 classifications in the VLA 3GHz survey and 155 in the Chandra survey. These were split into 812 star forming galaxies and 402 AGN. We also investigate cross matching with the MPA-JHU catalogue \citep{2003MNRAS.341...33K, 2004MNRAS.351.1151B, 2007ApJS..173..267S}, however, found that all matches were already represented by the VLA and Chandra surveys. Therefore, we only utilise these two catalogues in our sample.

While the ratio of AGN to star forming galaxies in our sample seems large compared to other works, it is important to note that this is a result of limited matching between the catalogues. Of the 4,182 galaxies in our sample, only 1,214 appeared at all in either the VLA or Chandra catalogues. Any galaxy which did not appear in these catalogues, we discard as unclassified. Therefore, we require these catalogues to specifically show us if a galaxy is star forming or not before we classify it as such.

\section{STAR FORMATION EVOLUTION WITH INTERACTION STAGE}\label{results}
\subsection{Controlling for Interaction Stage}
\noindent With our samples selected we investigate the change in multiple parameters with stage of the interaction. As stated previously, the four stages of interaction are designed to capture the main different parts of galaxy interaction. First, we show the results of breaking down our sample into stages with relation to the star formation and mass of our sample. We utilise the estimates of these parameters that exist in the COSMOS2020 catalogue itself. We have the option to use either the measured estimates from LePhare or EAzY. We opt to use the stellar mass measured from LePhare and the star formation rate measured with EAzY for our final results. We choose this as using any other combination reveals the degeneracys of the SED template fitting conducted in either of the underlying softwares. Figure \ref{fig:difference-measures} shows how the two measures differ from each other. As shown, EAzY measures larger stellar masses compared to LePhare while the scatter in SFR is very large.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/mass-sfr-scatter.pdf}
    \caption{Comparison of the measures of stellar mass and SFR using either LePhare or EAzY photometric codes to calculate them. \textit{Left}: The scatter in the stellar masses between softwares. As shown, EAzY often seems to find larger stellar masses when compared to LePhare. \textit{Right}: Scatter in SFRs when measured with LePhare or EAzY. There is a large difference between the measures of LePhare and EAzY here, and therefore we opt to only use EAzY measures of the SFR through this work.}
    \label{fig:difference-measures}
\end{figure}

As the measure of the masses between the softwares is similar, we can interchange which mass we use in these measurements. However, with the very large scatter of the SFRs between the softwares, we opt to only use the EAzY measurement. % Where do these differences come from? What potential impact could this have on our results?

With the measures of stellar mass and SFR selected, we break down our sample into their different stages. Figure \ref{fig:sfr-mass} shows the breakdown of these parameters with stage. On a population scale, there is clear evolution in the star formation rate from stage 1 through to stage 4. In stage 1, where the galaxies are distinct from one another with no clear morphological disturbance, we clearly see two populations of galaxies. These are the blue, star-forming and red sequences. The contours of number density in Figure \ref{fig:sfr-mass} bring out this population. In stage 2, when the galaxies are at pericentre, this red sequence remains but is highly diminished while the star forming sequence remains constant. Stage 3 shows a similar effect, where the red sequence reduces again before finally in stage 4 the red sequence completely disappears.

\begin{figure}
    \centering
    \includegraphics[width = 0.45\textwidth]{figures/sfr-mass-stages.pdf}
    \caption{The LePhare stellar mass against the EAzY SFR across the different stages of the interaction. The blue contours are 8 levels of density of the underlying populations in each frame. \textit{Top left}: The stellar mass and star formation rate of Stage 1 of our sample. Here, the interacting galxies are simply close pairs with little to no morphological disturbance. There are clearly two populations here: a main, star forming sequence forming the main population and a smaller red sequence. \textit{Top right}: Stage 2 of the interaction, where the two interactors are close to pericentre. The star forming sequence remains, but the red sequence is reduced significantly. \textit{Bottom left}: Stage 3 of the interaction, where the interactors are close to apocentre or escaped. Here, we see the almost complete disappearance of the red sequence. \textit{Bottom Right}: Stage 4 of the interaction, where the two systems are close to coalescence or have merged. The red sequence of galaxies has completely disappeared.}
    \label{fig:sfr-mass}
\end{figure}

However, this result could also be due to many other factors rather due to interaction stage. For instance, if the mass distribution of our sources evolves as well, it could just be that we are selecting higher mass systems in stage 4 when compared to stage 1. This would have the result of systems in stage 4 having, on average, higher SFRs than those in stage 1 and appearing like we had evolution in the star forming population with stage. Another effect that could cause this relation to appear would be our selection was highly dependent on galactic environment. It is well known that environment and star formation are closely linked, and that classifying interacting galaxies by morphology classifiers can weight up the cluster environment over galaxies in the field. In Section \ref{sec:env-cont}, we control for the environment in our sample and show that this is not an issue. Here, we investigate the question of the evolution of the mass distribution that could give us this result.

We can quantify the similarity of the mass distributions, and then the SFR distributions, using well known statistical tests. We opt to use two different tests: the KS-test and the Shapiro-Wilkes test. The KS-test is excellent at comparing different weighted distributions to each other, while the Shapiro-Wilkes test compares these distributions to other well known functions (in our case, the normal function). Thus, by weighting the mass distributions and using a KS-test, we will be able to quantify how alike the mass distributions are and then compare this to the weighted distribution of the SFRs. The weighting scheme we utilise normalises the distributions such that each bin in the distribution had the same number of sources within it. Therefore, any bins with few than the average number of sources will be weighted up while those bins with more will be weighted down. These weights applied to the mass distribution are then applied to the SFR distribution so the counts in each bin are, essentially, the same.

\begin{figure}
    \centering
    \includegraphics[width = 0.45\textwidth]{figures/stellar-mass-dist.pdf}
    \caption{The stellar mass distribution across the four stages. Each bin is weighted based on the counts in the smallest sub-sample in stage. This is stage 4 of interaction.}
    \label{fig:weighted-mass}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/sfr-dist.pdf}
    \caption{SFR distribution weighted by mass across each stage. This weighting is based on our sample of stage 4.}
    \label{fig:weighted-sfr}
\end{figure}

Figure \ref{fig:weighted-mass} shows the weighted mass distributions through the four different stages. These appear highly similar, and we conduct Kolmogorov-Smirnov \citep[KS-test;][]{an1933sulla} and Anderson-Darling \citep[AD-test;][]{stephens_74} tests to compare each distribution. In each test, we calculate the value of the KS test and a p-value which is the probability that each distribution is drawn from the same parent distribution. For each mass distribution, we find that p-value of the KS test is approximately one. Thus, it is not unreasonable to assume that the galaxies between each of our samples are drawn from the same parent sample and are very similar. Figure \ref{fig:weighted-sfr} shows the SFR distributions while being weighted and controlled for mass. When comparing stages 1 and 2, 1 and 3, 1 and 4, 2 and 4 and 3 and 4, the p-values are $\ll$0.05. Thus, we can reject the null hypothesis for these distributions and assume they are from different parent samples and are not identical. However, for comparing stages 2 and 3, the p-value of the KS test is $\approx$0.90. Thus, while these distributions are likely to be not identical, they are very similar in the parent sample they have been drawn from. \textbf{Thus, for the same mass distribution through each stage of interaction, the star formation distribution radically changes from stage 1 to stage 2 and from stage 3 to stage 4, while remaining reasonably similar from stage 2 to stage 3.}

Putting this result into the context of the dynamical timescale of an interaction, it shows there are three separate stages star formation evolution in these systems. The first is the change from isolated SFRs to interacting SFRs. Here, the SFR between galaxies which are close pairs and those that are at pericentre are radically different. This difference persists through to stage 3 - where the galaxies are at pericentre in the interaction. The SFR distribution remains approximately the same between these two, meaning the forces that drive and affect star formation remain equivalent between these two stages. Finally, the SFR radically changes again when we approach the merging or post-merger stage of the interaction. We can also say that this change is likely an enhancement in the SFR of the galaxies through the interaction due this change being driven by the disappearance of the red sequence through each stage.

We also control for redshift in our examination of this result. While the red sequence is significantly reduced across stage, it is important that we put this into the context of the star forming main sequence of galaxies. We expect that, due to an interaction, the population of starbursting and quiescent galaxies will change with stage. Therefore, we define a star forming main sequence through across redshift. This allows us to classify each source based on its SFR independently of redshift. To define the star forming main sequence (SFMS), we follow the example of \citet{2019MNRAS.484.4360A}. There, they define the star forming main sequence as a function of galactic stellar mass and redshift as

\begin{equation}
    \log \text{SFR}_{\text{MS}}(z)[M_\odot yr^{-1}] = -7.6 + 0.76\log\frac{M_*}{M_\odot} + 2.95\log(1+z).
\end{equation}

\noindent This finds the expected, main sequence, SFR of a galaxy at a given stellar mass, $M_*$, and redshift, z.

The ratio of the measured galaxy SFR and the expected main sequence SFR is then taken. We then use this fraction to classify each galaxy into distinct bins. These are:

\begin{enumerate}[(i)]
    \item Starburst galaxies. Here, the galaxy SFR is highly elevated compared to the SFMS. We follow \citet{2019MNRAS.484.4360A} and define a cutoff of $\log$SFR/SFR$_{\text{MS}} >$ 0.4.
    \item Main sequence galaxy. The galaxy is within 0.4 dex of the SFMS and approximately has the expected SFR. Defined as $-0.4 < \log$SFR/SFR$_{\text{MS}} < 0.4$.
    \item Sub-main sequence galaxy. A galaxy whose SFR is below the majority of the SFMS, but likely not quiescent. Defined as $-1.3 < \log$SFR/SFR $_{\text{MS}} < $ -0.4.
    \item Quiescent (High) galaxy. A galaxy with an SFR in the top $\approx$50\% of the quiscent galaxy population. Defined as $-2.3 < \log$SFR/SFR $_{\text{MS}} <$ -1.3.
    \item Quiescent (Low) galaxy. A galaxy with low SFR and very likely completely quenched. Defined as $\log$SFR/SFR $_{\text{MS}} <$ -2.3.
\end{enumerate}

\noindent We split our sample into its different stages and apply these criteria. Figure \ref{fig:sfr-clsf} shows the ratio between the expected SFMS SFR and the found SFR in COSMOS. This clearly shows a large increase in galaxies classified as starburst from stages 1 to 4 and a large reduction in the number of quenched systems. Figure \ref{fig:sfr-clsf-bar} shows the change in fraction of the different galaxy classifications through stage. These Figures reflect the results found in Figure \ref{fig:sfr-mass}. Initially, in stage 1, we find that the majority of our galaxies lie on the SFMS or just below it. There also exists a small population of galaxies which are classified as a starburst with a population of quiescent galaxies that is roughly double the fraction. As we move through the interaction stage, we see that the quiescent galaxy fraction gradually decreases to the point of almost non-existence in stage 4 galaxies. The complete inverse is true in our starburst fraction. We find that this almost quadruples in the fraction of our sample over the course of the different stages of interaction. The fraction of galaxies on the SFMS remains dominant throughout, however, we do find the fraction of sub-MS galaxies significantly reduced.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/sfr-clsf-dist.pdf}
    \caption{Stellar mass against the ratio of measured SFR to the expected SFR if the galaxy was on the SFMS. Black points are the individual sources, while the blue contours are as in Figure \ref{fig:sfr-mass}. The red dotted lines show the cutoffs for different galaxy classifications based on their SFR, with each cut off being defined by the text in blue. We find that through interaction stage, the quiescent galaxy population significantly reduces while the starburst population rapidly increases. As these cutoffs are also dependent on redshift, we find that this evolution in SFR with interaction stage is independent of redshift.}
    \label{fig:sfr-clsf}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/sfr-clsf-bar.pdf}
    \caption{The change in fraction of different galaxy classifications from the fraction of SFR to the expected SFR on the SFMS. While galaxies on the SFMS remain dominant through each sample across interaction stage, there is significant change in the starburst and quiescent populations. The starburst galaxy population moves from being roughly half the size of the quiescent population in stage 1 to completely dominating it in stage 4. This is occurring while the quiescent population is significantly reduced to almost non-existence in stage 4.}
    \label{fig:sfr-clsf-bar}
\end{figure}

% So, what does this all mean?
Thus, we find that throughout the different stages of interaction the star formation within the systems is being driven upwards. We have found here is that this process of increasing SFR and galaxies being driven to starburst begins from the pericentre of the first interaction. Simulations have long showed that there is a peak in SFR initially in an interaction, followed by a slow reduction to a final enhanced level in the post-merger stage. However, due to the initial starburst, it is often shown that the gas within the galaxy is either quickly used up and the galaxy quenches or it is heated to the point that star formation becomes very inefficient. Thus, what we would expect to see here is that stages 2 and 4 would have the highest fractions of galaxies undergoing a starburst and the smallest quiescent fractions.

Our work here is found to be correct by both simulations and previous works on when the starburst occurs in an interacting galaxy. We observe significant change in the starburst and quiescent fractions from stage 1 to 2 and stage 3 to 4. Thus, the initial part of the interaction appears to have a significant impact on the SFR (and, therefore, the gas content) of the galaxies involved. This enhanced star formation then persists through the interaction, even when the other galaxy approaches the apocentre of the interaction. This is seen by the relatively little change in the SFR classifications from stage 2 to 3. While the fraction of MS galaxies does increase slightly, the fraction of starburst galaxies remains essentially constant. There is also no major change in the quiescent fraction either. However, it is interesting to note that the sub-MS fraction does decrease. Thus, the transition to stage 2 to stage 3 of an interaction may bring with it slight enhancement in the SFR, but not on the scale of the other transitions in interaction stage. We then see clear evidence for a second starburst is being driven by the merging of the systems that goes beyond even the enhanced from the initial pericentre passage.

So, what does the SFR evolution look like from our results? As the interactors approach each other, they are representative of a control sample where there is a population of quiescent, MS and starburst galaxies. However, upon transitioning from stage 1 to 2, the disturbance due to interaction provokes increased fragmentation of a galaxies gas and leads to an increase in star formation. This is true even of the galaxies which have very little gas remaining, as the quiescent fraction reduces significantly in our sample. This enhancement persists from stage 2 to stage 3, where we see no major change in the classification fractions. Then, upon merging and transitioning to stage 4, we find that even the most gas-empty galaxies begin forming stars quickly.

Now, it is important to note the parameter space that we are searching in these examples. We are probing major interactions between galaxies, where the resultant tidal features that form would be classifiable in an image. Hence, even those mergers which are considered dry - with very little gas - would be capable of driving major rejuvenation in the star formation of these systems. To conclude, \textbf{we find clear, statistically robust evolution of a galaxys' SFR with interaction stage. The SFR changes dramatically from stage 1 to 2 and stage 3 to 4 - at the pericentre and coalescence of the interaction. Existing simulations of galaxy interactions support this interpretation of when star formation changes.}

\subsection{Projected Separation and Star Formation Enhancement}
\noindent To further confirm this result, we directly investigate the relation between the star formation enhancement (SFE) and the projected separation using our confirmed sub-sample of galaxy pairs. This sample is significantly smaller than our non-pair sample: containing 480 pairs or 960 galaxies. Upon sub-dividing this into different stages we find 310 stage 1 galaxies, 146 stage 2 galaxies and 498 stage 3 galaxies. Only 3 stage 4 galaxies were in our confirmed galaxy pair sub-sample, and therefore we do not attempt to make inferences about this population.

To measure the SFE of our galaxy pairs, we directly compare the to the control sample that was also created with this sub-sample and defined in Section \ref{sec:sec-ident}. We separate our galaxy pair sample into different bins based on the projected separation between them. We find that the bulk of our sample has a projected separation between the two galaxies of $\leq$50kpc. Therefore, we sample from this region of the parameter space with high precision and smaller bin widths before we increase the bin widths at larger projected separations. We define a cutoff that each bin must contain at least 10 counts to be included in this plot and, therefore, by changing the bin widths with projected separation we are able to maintain some level of statistical robustness. We define our bins as [0.5, 10, 20, 50, 100, 125, 150] kpc.

We create distributions based on these bins of projected separation: one of interacting galaxies and the other of our control galaxies. We then take the average of each bin separately, before dividing each distribution by the other. By taking the average of each bin, we are finding the total excess of the SFR caused by interaction alone and then comparing it to what we would expect from non-interacting galaxies in the same bin. Note, that the control galaxy relation to the projected separation is meaningless as they are not paired. The control galaxies are simply mass and redshift matched to each interacting galaxy and used as a reference for what SFR we would expect of it had it not been interacting. Finally, we divide the the averaged interacting SFR distribution by the averaged control SFR distribution. This provides us with a measure of the excess star formation due to interaction, and the SFE when compared to the regular population.

\begin{figure*}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/sfr-enhancement-projected-sep.pdf}
    \caption{The projected separation against the star formation enhancement in average star formation at different bins of projected separation. Each bin must contain at least 10 counts to be considered. As the bulk of our galaxy pair sample is at low projected separation, we heavily sample from this region of the parameter space. The bins are: [0.5, 10, 20, 50, 100, 125] kpc. As we move to higher projected separation, the bins increase in width to maintain statistical significance in our sample. \textit{Left}: The star formation enhancement found in the entire galaxy pair sample. As expected, we see a gradually decreasing enhancement. \textit{Right}: As left but broken up into different stages of the interaction. Black markers are stage 1, red stage 2 and blue stage 3. We only investigate stages 1,2 and 3 as only three galaxy pairs were identified in stage 4. Here, we find a generally decreasing star formation enhancement with projected separation but very different individual behaviour dependent on the stage classification.}
    \label{fig:sfr-enhancement-sep}
\end{figure*}

Figure \ref{fig:sfr-enhancement-sep} shows star formation enhancement between our interacting and control binned SFRs and the projected separations between each galaxy pair. On the left of the plot, we have the distribution for the full galaxy pair sample, without taking account of stage. The error bars are calculated based on the counts within the bin and errors on the underlying SFRs of the galaxies within. The dashed black line represents no enhancement in SFR, as the average SFR in the interacting galaxy sample would be equal to the average SFR in the control sample. We find at very small projected separation a SFR enhancement of 1.87 which gradually decreases with projected separation down to approximately 1 at 117.5kpc. On the right, we break the galaxy pair sample into different stages and plot out the resultant enhancement in star formation. This shows very different behaviour in enhancement dependent on the stage of the interaction.

In stage 1, we find that the star formation is very weakly enhanced, at a level of approximately 1.25. As we increase the projected separation bins, the enhancement then changes around an enhancement of 1, even dropping below 1 at 40kpc. Taking into account the errors on our measurements, the true value of the enhancement could be around 1 for the entire projected separation distribution. This is not unexpected. Stage 1 represents when the two galaxies are close pairs, but with no morphological disturbance. Therefore, we would not expect significant, if any, enhancement in the SFR based on the projected separation. This is the stage with the lowest enhancement across projected separation, remaining close to 1.

In stage 2, we see consistent enhancement with projected separation. It is also much larger than stage 1, with it being approximately 1.75 up to 1.90. In fact, the measured enhancement increases across the 50kpc of projected separation. However, accounting for the errors on these measurements and the declining counts in each bin, it is likely that the enhancement does drop as the projected separation increases. We see a very large enhancement in the overlapping stage 3 galaxies which overlap with the stage 2 measurements. The enhancement in stage 3 galaxies then rapidly declines as we move to higher projected separations. After 150kpc, our galaxy pair sample does not have the counts to make robust estimates of the SFE.

Finding that the SFE declines with projected separation is not an unexpected. Previous works also find this and even match our results \citep{2000ApJ...530..660B, 2008AJ....135.1877E, 2013MNRAS.433L..59P}. Where our result differs is by taking into account the stages of the interaction. \textbf{While using only projected separation is a proxy for stage, we find the actual SFE is also dependent on the time of in the dynamical timescale the interaction is}. For instance, all of our stage 2 systems have high SFE of greater than 1.5 no matter the projected separation between them. This shows that the morphological connection and overlap of the two systems is driving the enhancement. Then, moving to stage 3 with low projected separation, we see the SFE grows even larger to 2.5. The galaxy pair will be just at the point of breaking overlap and morphological connection here, only just moving away from pericentre here. Simulations have predicted that once the two systems are just passed pericentre and moving away from each other, the two galaxies will undergo an intense delayed starburst \citep{2015MNRAS.448.1107M}. We see this effect in the sudden increase in SFE as the two galaxies disconnect from each other and begin to move away from one another.

These distributions of SFE paint a picture of the disturbance and change in the SFRs of these systems throughout the interaction, we will summarise here. First, in stage 1, the two systems move towards one another (or away with little disturbance) with no SFE and matching the control. Upon entering into stage 2, the galaxies SFR dramatically increases due to the morphological disturbance of the two disks and potential overlapping and linking of the two systems. Then, once the two systems are past pericentre and enter into stage 3 to flyby one another, the gravitational torques on the two systems take hold and star formation is, again, dramatically increased. This increase relaxes as the systems move towards apocentre and higher projected separations.

However, it is important to note that this result has been found with a sample of 960 interacting galaxies in their pairs. No stage 4 galaxies have been represented and in numerous projected separation bins the counts are small enough to have conclusion altering error bars. \textbf{Therefore, it is imperative that future works, with larger sample sizes not only look at the projected separation of the systems but the morphology as well.}

\subsection{Controlling for Galactic Environment} \label{sec:env-cont}
\noindent While we have found clear evidence of evolution in the SFR of interacting galaxies with stage, it is important we remove a potential source of contamination. It is well known that the environment has a direct impact on the observed SFR of galaxies. A galaxy in a cluster environment has, on average, a higher SFR than those in the field \citep{2006MNRAS.373..469B}. Thus, if any of our stage classifications are biased towards one environment or another, it could have severe consequences for our results. Therefore, we utilise the existing COSMOS environment catalogue described in \citet{2017ApJ...837...16D} to ensure there are no environmental biases in our staging.

Upon applying the matching described in Section \ref{data:environ}, we break our samples into different environment measures with stage. Figure \ref{fig:dens-stage} shows the distribution of our matched samples with their density values. It is important to note that \citet{2017ApJ...837...16D} has a higher mass cutoff than we have implemented in our underlying sample. Therefore, this is showing the density of all systems $\log$M$_*$/ M$_\odot \geq$ 9.6.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/density-stage.pdf}
    \caption{The density about each of our sources matched with the \citet{2017ApJ...837...16D} catalogue. As shown, there is no existing bias in the distribution of galactic environments throughout our stages. This is confirmed by performing KS- and AD-tests. Therefore, we can conclude that the shown evolution in SFR with interaction stage is not due to environmental effects.}
    \label{fig:dens-stage}
\end{figure}

We conduct weighted KS and AD tests on the environment distributions, and confirm they are likely drawn from the same parent sample. The resultant p-values are $\approx$1 when comparing between the distributions. Thus, there are no biases in our stage selection with environment, and it fair to assume that the evolution in the SFR with stage is due to interaction, not environment.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/sfr-mass-density.pdf}
    \caption{The distribution of environment classifications through our sample. This is only for sources with a $\log$M$_*$ / M$_\odot \geq$ 9.6. However, from this subsample, we see that there is no trend with environment and that the distribution is random throughout each stage. Therefore, the effects of increasing SFR with interaction stage is not due to environmental effects.}
    \label{fig:dens-sfr-mass}
\end{figure}

Figure \ref{fig:dens-sfr-mass} shows the environment classification in our reduced stellar mass-SFR distribution. This shows that there is no ordered structure, or bias, in our sample with environment and further reinforces that the galactic environment is not responsible for our proposed SFR evolution with stage. The majority of our sample lie either within filaments or in the field. The environment which would have the most effect upon the SFRs of our sample is a cluster environment. However, the fraction of our sample within a cluster is never greater than 10\%. It is important to note, however, that the galaxies most likely to lie in a cluster in our sample would be lower mass galaxies. However, these systems are not found in the red sequence where we see the most change in SFR due to interaction. Therefore, the lack of information here does not have a large impact on this measurement.

\section{Nuclear Activity with Interaction Stage}
\noindent We now cross match our sample with the two AGN catalogues and methods described in Section \ref{sec:agn-clsf}. This provides us a sample of 1,361 AGN and star forming galaxies (SFGs). The breakdown of number of classifications per stage is shown in Table \ref{tab:agn-sfg-breakdown}. Figure \ref{fig:agn-stage} shows distribution of stellar mass to SFR with stage, with the confirmed AGN and SFGs marked. We find no major changes in AGN with stellar mass or SFR with stage in our sample. Galaxies containing AGN appearing throughout the SFMS and red sequence across each stage. There is no obvious bias in stellar mass or SFR for hosting an AGN. We do, however, see a concentration of AGN in stage 1 on the red sequence at higher masses. This shifts as we increase interaction stage, with the bulk of the AGN beginning to appear at lower masses in stages 3 and 4. Finally, in stage 4 we see AGN and SFGs distributed almost evenly across the star forming main sequence with little bias in distribution with mass. This hints at evolution with stage.

Applying KS- or AD- tests these, however, reveals values close to 1 and therefore it is very likely that these distributions have been drawn from the same parent sample. However, due to the low counts of AGN and SFGs in our sample, using KS- and AD-tests to verify this is not optimal. Therefore, we instead investigate the global AGN fractions in our sample across different stages. We apply a weighting, as in our SFR and stellar mass distribution examples, to account for the various counts across the different stages. We then take a second weighting of the measured AGN fraction from the total size of the different samples we have. Thus, we have assumed that we have equivalent counts in each mass bin of our sample as well as approximated having the same sample size in each stage bin.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/agn-stage-dist.pdf}
    \caption{The distribution of AGN through stage with SFR and stellar mass. We do not find any major changes and evolution with AGN and SFG position through interaction stage. }
    \label{fig:agn-stage}
\end{figure}

\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
         & Stage 1 & Stage 2 & Stage 3 & Stage 4 \\
         \hline
        SFG & 135 & 212 & 337 & 199 \\
        AGN & 103 & 117 & 163 & 95
    \end{tabular}
    \caption{Breakdown in number of classified AGN and SFGs per stage. With our counts so low in this sample, it is difficult to make concrete conclusions about the evolution of AGN during interaction.}
    \label{tab:agn-sfg-breakdown}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/agn-frac-time.pdf}
    \caption{The change in AGN fraction with stage. We define the AGN fraction as the ratio of number of confirmed AGN divided by the number confirmed AGN and SFGs in the sample. Errors on each point are found by bootstrapping. We do not include any sources of which we have no information. We find a sudden drop in AGN fraction from stage 1 to stage 2, followed by a rapid increase from stage 3 to 4.}
    \label{fig:agn-frac-time}
\end{figure}

Figure \ref{fig:agn-frac-time} shows the changing AGN fraction with interaction stage. We find that the AGN fraction generally increases with interaction stage. There is a large increase from stage 1 to 2, followed by the fraction mostly remaining unchanged until stage 3. Then, in stage 4 we measure another large jump in the AGN fraction. However, before we discuss this result further, we must point out that the large error bars on our measurements and the low number counts of confirmed AGN and SFG make this difficult to interpret. We measure the errors on our sample from bootstrapping the underlying low-count population of AGN and SFGs and comparing to the change in fraction measured. Thus, while our inital values show a general increase, because of the large error bars, there could be no increase at all with stage to even a slight decline.

However, finding that AGN fraction increases with stage is inline with other works \citep{2011MNRAS.418.2043E, 2023ApJ...942..107S} which often find an increase in AGN fraction with projected separation. Stage can also be considered as a proxy for the projected separation of the two systems, where stage 4 will be systems with the smallest projected separation and stage 3 being the largest separation. This matches what we find here, with stage 4 systems having the highest fraction of AGN compared to stage 1 systems. What is different in our results, however, is the change in AGN fraction from stage 2 to 3. While we see a large increase in AGN fraction from stage 1 to 2, this enhancement holds from stage 2 to 3. Again, this is not an unusual result and works such as \citet{2011MNRAS.418.2043E} show that the enhancement of nuclear ignition can actually be delayed in galaxy interaction. Thus, we only see the final ignition in the later part of the dynamical time than the initial flyby.

% Section here on actually looking at the projected separation of the two systems.
It is also difficult to use interaction stage as a proxy for the projected separation in this context. As shown previously, stages 2 and 3 have some overlap in the range of projected separations we have classified them into. Therefore, we also investigate any overlap between our confirmed pairs and the AGN fractions we see here. We find 104 AGN and 180 SFGs overlap with our confirmed pairs. This sample is far too small to breakdown further into different stages. Therefore, Figure \ref{fig:sfg-agn-proj} shows the change in density of AGN classifications and SFGs with increasing projected separation. As expected, we find that with increasing projected separation the number of confirmed AGN and SFGs decreases. However, in the AGN distribution, we find two contributing components in it. The first is a peak in the projected separation distribution is from 0kpc-25kpc. The only contribution here is from stage 2 galaxies, which are close to pericentre. The second peak, from a projected separation of 30kpc - 60kpc is dominated by stage 3 galaxies, although some stage 1 AGN are in this peak as well. Finally, the third peak from 85 - 125kpc is representative of a mixture of stage 1 and stage 3 galaxies. None of our stage 4 galaxies which had their secondaries identified overlapped with our AGN and SFG identified sample.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/sfg-agn-dist.pdf}
    \caption{The distribution of both AGN and SFGs which have been cross matched with our confirmed galaxy pairs. While we find, as expected, the number of AGN and SFGs decreases with projected separation we find interesting structure within the AGN sample. Here, there appears to be a double peak in the number of counts of AGN. The first peak, between a projected separation of 0 and 25kpc is primarily from stage 2 galaxies while the second peak of 40 to 65kpc is of stage 3 galaxies in our sample. As shown in Figure \ref{fig:agn-frac-time}, we expect the fraction of AGN to be similar in these two stages while they are at very different times in the dynamical timescale of the interaction. The third peak, from 75 to 125kpc is only from stage 1 and stage 3 galaxies.}
    \label{fig:sfg-agn-proj}
\end{figure}

Thus, the observed enhancement in AGN across a range of different projected separations if primarily due to different parts of the dynamical time of an interaction. The peak in nuclear activity, as expected, is within stage 2 galaxies which are close to pericentre in their interaction. However, as we move to higher projected separations and towards stage 3 of the interaction, the number of AGN begins to decrease. This implies, following other works, that enhancement occurs early in the interaction, when the systems are at pericentre from each other and with the smaller the pericentre, the more likely it is of nuclear ignition. The nuclear ignition from stage 2 can then be switched off again as we move to higher projected separation and stage 3. Finally, as the interaction begins to move into coalescence, we find another large increase in the AGN fraction which indicates that AGN activity becomes enhanced. Therefore, \textbf{when finding the AGN fraction of interacting systems it is important to consider the stage, and relate this to the projected separation of the system.}

\section{Discussion}
\subsection{Interaction Stage and Projected Separation}

\subsection{Interaction Stage and AGN}

\subsection{Comparison to Simulations}

\section{Conclusions}\label{conclusion}
 - Restate the Results \\
 - What's the next steps of this project?
    - Hint at what I would do in a Fellowship.

\section*{Acknowledgements}
The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

This work utilised data from the Cosmic Evolution Survey (COSMOS) which has a digital object identifier of 10.26131/IRSA178.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:
\bibliographystyle{mnras}
\bibliography{references} % if your bibtex file is called example.bib

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Needed?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex